On the Gravitational stability of gravito-turbulent disks

### 2015-12-22 paper/appendix3d.tex
re-doing appendix B

\section{2D viscous GI} \label{gammie_check}
We recover \cite{gammie96}'s dispersion relation 
for viscous gravitational instability as follows. First we set
$\mu=-1, \lambda=0$ to obtain the same viscosity models. Next we
simplify Eq. \ref{bigA}---\ref{bigF} by assuming $|\beta S |\gg 1$  
%This will turn out to be equivalent to $\beta\gg 1$, i.e. the
%adiabatic limit. 
and $S\lesssim O(K^2)$ as $|K|\to 0$. Applying these considerations to
Eq. \ref{thindisk} gives 
\begin{align}
  S = \frac{\alpha K^2\left[\frac{2|K|}{Q} - \gamma K^2 +
      2q(2-q)(\gamma-1)\right]}{2(2-q) + \gamma K^2 - \frac{2K}{Q}}, 
\end{align}
which does not explicitly depend on cooling. 
Setting $\gamma=1$ then gives \citeauthor{gammie96}'s Eq. 18. The above
is expression thus generalizes \citeauthor{gammie96}'s result to
non-isothermal disks.

>>>>

\section{2D viscous GI} \label{gammie_check}
We obtain the dispersion relation for viscous GI described by
previous authors \citep{lynden-bell74,willerding92,gammie96} as 
follows. We set $\mu=-1, \lambda=0$ to obtain the same viscosity
models. Next we consider $|\beta S|\to \infty$, i.e. no cooling. Then
the condition $AD = BC$ imply
\begin{align}
  S^3 + \left(\frac{7}{3}\alpha + \alpha_b\right)K^2S + \left[2(2-q) -
    \frac{2|K|}{Q} + \gamma K^2 + \alpha K^4 \left(\frac{4}{3}\alpha +
    \alpha_b\right)\right] + \alpha K^2 \left[\gamma K^2 -
    \frac{2|K|}{Q} - 2q(2-q)(\gamma-1)\right]=0,
\end{align}
which agrees with \citet[Eq. 11]{willerding92} in the isothermal
limit ($\gamma=1$). The non-isothermal term $\propto (\gamma-1)$
originates from viscous dissipation which was excluded in the
aforementioned works. Its effect is to increase 
(decrease) the maximum wavenumber (wavelength) allowed for viscous
GI.

An

%We recover \cite{gammie96}'s dispersion relation 
%for viscous gravitational instability as follows.  Next we
%simplify Eq. \ref{bigA}---\ref{bigF} by assuming $|\beta S |\gg 1$  
%This will turn out to be equivalent to $\beta\gg 1$, i.e. the
%adiabatic limit. 
%and $S\lesssim O(K^2)$ as $|K|\to 0$. Applying these considerations to
%Eq. \ref{thindisk} gives 
%\begin{align}
%  S = \frac{\alpha K^2\left[\frac{2|K|}{Q} - \gamma K^2 +
%      2q(2-q)(\gamma-1)\right]}{2(2-q) + \gamma K^2 - \frac{2K}{Q}}, 
%\end{align}
%which does not explicitly depend on cooling. 
%Setting $\gamma=1$ then gives \citeauthor{gammie96}'s Eq. 18. The above
%is expression thus generalizes \citeauthor{gammie96}'s result to
%non-isothermal disks.

### 2015-12-22 paper/paper3d.tex
re-doing appendix B

\subsection{Viscosity removes rotational support}
It is well known that viscous, self-gravitating disks can develop 
GI even when $Q>1$. This is because viscosity compromises rotational
support against self-gravity for long-wavelength disturbances
\citep{lynden-bell74,willerding92,gammie96}.

@@

How well turbulence can be modeled as an effective, Navier-Stokes
viscosity is a separate issue \citep{balbus99}. However, it is 
reasonable to speculate that turbulence frustates rotational 
support. (This is, after all, a physical cause for mass accretion.)

Again, viscosity is absent in the Toomre $Q$ parameter. The above
studies account for viscosity but consider isothermal disks. On the
other hand, proper inclusion of visosity should also account for
viscous heating. We thus extend the study of viscous GI to 
include an energy equation with viscous heating and explicit
cooling. This allows us to consider GI in realistic PPDs.

@@

\subsection{Relation to gravito-turbulent disk fragmentation}
The two effects described above are immediately relevant to
`gravito-turbulence'. This is a steady disk state in which cooling is
balanced by turbulent heating, itself caused by GI. In this case 
the $\alpha$ and $\beta$ parameters are inversely related.

@@

>>>>

\subsection{Viscosity removes rotational support}
It is well known that viscous, self-gravitating disks can develop 
GI even when $Q>1$. This is because viscosity compromises rotational
support against self-gravity for long-wavelength disturbances
\citep{lynden-bell74,willerding92,gammie96}. %also in dusty fluids -
                                %youdin 11 and the inutsuka paper

@@

How well turbulence can be modeled as an effective, Navier-Stokes
viscosity is a separate issue \citep{balbus99}. However, it is 
reasonable to assume that turbulence frustates rotational 
support. (This is, after all, a physical cause for mass accretion.)

Again, viscosity is absent in the Toomre $Q$ parameter. The above
studies account for viscosity but consider isothermal disks. On the
other hand, proper inclusion of viscosity should also account for
viscous heating which, as discussed above, also affect stability. 
We thus extend the study of viscous GI to 
include an energy equation with viscous heating and explicit
cooling. This allows us to consider GI in realistic PPDs.

@@

\subsection{Relation to gravito-turbulent disk fragmentation}%and ring
                                %formation 
%can the continous formation/destruction of clumps (as seen in sims)
%be interpreted as `overstability'? 
%for overstable modes over-density oscilates 
The two effects described above are immediately relevant to
`gravito-turbulence'. This is a steady disk state in which cooling is
balanced by turbulent heating, itself caused by GI. In this case 
the $\alpha$ and $\beta$ parameters are inversely related.

@@

%viscous gi speculated by gammie 01 but now found in those sims. but
%longer sims of the same system by paardekooper eventually fragment
